---
title: "Analyzing larval environmental data - Moorea 2023"
author: "Ariana S Huffmyer"
date: "2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo tidbit loggers and daily measurements.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Daily measurements  

## Read in file  

Read in daily measurements file.  

```{r}
daily<-read_csv("data/environmental/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tank<-as.factor(daily$tank)
daily$tris.date<-as.character(daily$tris.date)
```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/environmental/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*Ttris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = pH(Ex=pH.mV, Etris=mVTris, S=sal.psu, T=temp.C)) # calculate pH of the tanks using the pH seacarb function
```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  
```{r}
daily_calc<-pHSlope%>%
  select(date, time, tank, temp.C, pH.total, sal.psu, par)
```

## Change to long format

Change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:par,
  names_to = "metric",
  values_to = "value")
```

Filter by relevant dates. 
```{r}
#daily_calc.long<-daily_calc.long%>%
#  filter(date>as.POSIXct("2023-06-18"))
```

## Plot metrics of interest  

Plot by tank colored by tank.      
```{r}
daily_tanks<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=tank))+
  geom_point(size=2, position=position_dodge(0.3))+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_classic(); daily_tanks
  
ggsave(filename="figures/environmental/tank_daily_measurements.png", plot=daily_tanks, dpi=300, width=6, height=8, units="in")

```

## Test for differences between tanks  

Obtain a short format dataset.  
```{r}
daily_calc.short<-daily_calc.long%>%
  spread(metric, value)
```

Test for differences between tanks. 

```{r}
daily_calc.short%>%
  aov(par~tank, data=.)%>%
  summary()

daily_calc.short%>%
  aov(pH.total~tank, data=.)%>%
  summary()

daily_calc.short%>%
  aov(sal.psu~tank, data=.)%>%
  summary()

daily_calc.short%>%
  aov(temp.C~tank, data=.)%>%
  summary()
```

Output statistical results to a table. 
```{r}
stats<-daily_calc.long
  
metric_list <- unique(stats$metric)

# Create an empty data frame to store the ANOVA results
anova_results <- data.frame()

# Loop through each metric and perform one-way ANOVA
for (i in metric_list) {
  anova_result <- stats %>%
    filter(metric == i) %>%
    aov(value ~ tank, data = .) %>%
    tidy()

  # Add the metric column to the ANOVA result
  anova_result$metric <- i

  # Append the ANOVA result to the anova_results data frame
  anova_results <- rbind(anova_results, anova_result)
}

# Replace "anova_results.csv" with the desired filename
write.csv(anova_results, file = "output/environmental/anova_results.csv", row.names = FALSE)
```

## Summarize daily measurements  

Calculate descriptive statistics   
```{r}
summary<-daily_calc.short%>%
  group_by(tank)%>%
  select(!date)%>%
  select(!time)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

write_csv(summary, "output/environmental/daily_measurements_summary.csv")
```
