---
title: "Protein analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")

# load packages
library(tidyverse)
library(broom)
```

# Import data
```{r}
# List protein data files
prot_path = "data/physiology/protein/"                                              # Path to prot data directory

#/Users/hputnam/MyProjects/moorea_symbiotic_exchange_2023/data/physiology/protein/20231207_Plate1_BCA_platemap.csv

all_prot_files <- list.files(path = prot_path, pattern = "*.csv")          # List all files in directory
prot_platemaps <- list.files(path = prot_path, pattern = "platemap")       # List platemap files
prot_data_files <- setdiff(all_prot_files, prot_platemaps)                 # List data files

# Read in all files into tibble
df <- tibble(file = prot_data_files) %>%
  separate(file, into = c( "date", "plate"), remove = FALSE) %>%
  unite(plate, date, plate) %>%
  mutate(platemap = map(plate, ~read_csv(paste0(prot_path, ., "_BCA_platemap.csv"))),
         prot_data = map(file, ~read_csv(paste0(prot_path, .)) ))

# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, prot_data, ~ right_join(.x, .y)))
```

# Plot standard curve
```{r}
# Create standard curve following kit instructions
standards <- tribble(
  ~std, ~BSA_ug.mL,
  "A",        2000,
  "B",        1500,
  "C",        1000,
  "D",         750,
  "E",         500,
  "F",         250,
  "G",         125,
  "H",          25,
  "I",           0
)

std_curve <- df %>%
  unnest(merged) %>%
  filter(grepl("Standard", colony_id)) %>%
  select(plate, Well, colony_id, abs562 = `562:562`) %>%
  rename(std = colony_id) %>%
  mutate(std = str_sub(std, 9, 9)) %>%
  #group_by(std) %>%
  #summarise(abs562 = mean(abs562)) %>%                       # calculate mean of standard duplicates
  #mutate(abs562.adj = abs562 - abs562[std == "I"]) %>%       # subtract blank absorbace value from all
  left_join(standards)


plot(std_curve$BSA_ug.mL, std_curve$abs562)
std_curve

std_curve <- std_curve %>%
  filter(Well!="G8") 

## Fit linear model for standard curve
 mod <- lm(BSA_ug.mL ~ abs562, data = std_curve)
# coef(mod)

## Fit nonlinear model for standard curve
#mod <- nls(formula = BSA_ug.mL ~ z + a * exp(b * abs562), start = list(z = 0, a = 1, b = 1), data = std_curve)

fitted <- mod %>% broom::augment()

# Plot standard curve
std_curve_plot <- std_curve %>%
  ggplot(aes(x = abs562, y = BSA_ug.mL)) +
  geom_point(color = "red", size = 3) 

std_curve_plot + 
  geom_line(data = fitted, aes(x = abs562, y = .fitted)) +
  labs(title = "Standard curve")
```

# Calculate protein concentrations
```{r}
# Calculate protein concentration for all samples using standard curve
prot <- df %>%
  unnest(merged) %>%
  filter(!grepl("Standard", colony_id)) %>%                     # Get just samples (not standards)
  select(plate, Well, colony_id, abs562 = `562:562`) %>%        # Select only needed columns
  filter(!is.na(colony_id)) %>%                                 # Filter out empty wells
  filter(colony_id != "BK") %>%                                 # Filter out blank wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod, newdata = data.frame(abs562 = .))))    # Use standard curve to convert absorbance to protein

std_curve_plot + 
  geom_point(data = prot, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve")
```

# Normalize to surface area
```{r}
# Surface area data
sa <- read.csv("output/physiology/surface_area/surface_area.csv")
colnames(sa)[1] <- "colony_id"

# Tissue homogenate volume data
homog_vols <- read_csv("data/physiology/homog_vols.csv") 
colnames(homog_vols)[1] <- "colony_id"

# Coral sample metadata
metadata <- read_csv("data/field_collection/collection_metadata.csv") 
colnames(metadata)[1] <- "colony_id"

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols, by="colony_id") %>%
  full_join(sa, by="colony_id")

# Join prot data with metadata
prot <- left_join(prot, metadata) %>%
  mutate(prot_ug = prot_ug.mL * total_slurry_ml,
         prot_ug.cm2 = prot_ug / surface.area.cm2,
         prot_mg.cm2 = prot_ug.cm2 / 1000)
```

# Plot results by species and stage
```{r}
# Plot all data points with mean ± se
prot %>%
  filter(!is.na(species)) %>%
  ggplot(aes(x = lifestage, y = prot_mg.cm2, color = species)) +
  facet_wrap(~species)  +
  coord_cartesian(ylim = c(0, 1.75))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean

# Remove outlier data points where absorbance was super low
#prot <- prot %>%
#  filter(abs562 > 0.1)
```

# Write data to output file
```{r}
# Write protein data to output file
prot <- prot %>%
  group_by(colony_id,species,lifestage) %>%
  summarise(prot_mg.cm2 = mean(prot_mg.cm2, na.rm = T)) %>%
  select(colony_id, species, lifestage, prot_mg.cm2) 

  write_csv(prot, file = "output/physiology/protein/protein.csv")
```

# Plot results by species and stage
```{r}
# Plot all data points with mean ± se
prot %>%
  filter(!is.na(species)) %>%
  ggplot(aes(x = lifestage, y = prot_mg.cm2, color = species)) +
  facet_wrap(~species)  +
  coord_cartesian(ylim = c(0, 1.75))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean

# Remove outlier data points where absorbance was super low
#prot <- prot %>%
#  filter(abs562 > 0.1)
```

# Plotting and ANOVA analyses 

Attach temperature metadata. 
```{r}
sample.info <- read_csv(file = "data/p_r_rates/pr_sample_metadata.csv")
sample.info<-sample.info%>%rename(colony_id=sample_id)%>%select(colony_id, Temp.Cat)%>%filter(!grepl("BK", colony_id))

prot<-left_join(prot, sample.info)
```

View individual observations. 
```{r}
prot_plot1<-prot%>%
  
  ggplot(aes(x=Temp.Cat, y=prot_mg.cm2, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot1

ggsave("figures/physiology/protein/protein_dots.png", prot_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
prot_plot2<-prot%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(prot_mg.cm2, na.rm=TRUE), sd=sd(prot_mg.cm2, na.rm=TRUE), N=length(prot_mg.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot2

ggsave("figures/physiology/protein/protein_means.png", prot_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature.  
```{r}
prot_plot3<-prot%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(prot_mg.cm2, na.rm=TRUE), sd=sd(prot_mg.cm2, na.rm=TRUE), N=length(prot_mg.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot3

ggsave("figures/physiology/protein/protein_overall_means.png", prot_plot3, width = 6, height = 4)
```

## Run ANOVA model

```{r}
model1<-prot%>%
  
  aov(prot_mg.cm2~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Strong effects of species and lifestage, no interactions. In POC and POR, adult protein is higher than recruit protein. No difference in ACR detected. There is a difference in ACR-POR and POC-POR. No difference between ACR and POC at a species level.  



