---
title: "Protein analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")

# load packages
library(tidyverse)
library(broom)
library(emmeans)
```

# Import data
```{r}
# List protein data files
prot_path = "data/physiology/protein/"                                              # Path to prot data directory

#/Users/hputnam/MyProjects/moorea_symbiotic_exchange_2023/data/physiology/protein/20231207_Plate1_BCA_platemap.csv

all_prot_files <- list.files(path = prot_path, pattern = "*.csv")          # List all files in directory
prot_platemaps <- list.files(path = prot_path, pattern = "platemap")       # List platemap files
prot_data_files <- setdiff(all_prot_files, prot_platemaps)                 # List data files

# Read in all files into tibble
df <- tibble(file = prot_data_files) %>%
  separate(file, into = c( "date", "plate"), remove = FALSE) %>%
  unite(plate, date, plate) %>%
  mutate(platemap = map(plate, ~read_csv(paste0(prot_path, ., "_BCA_platemap.csv"))),
         prot_data = map(file, ~read_csv(paste0(prot_path, .)) ))

# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, prot_data, ~ right_join(.x, .y)))
```

# Plot standard curve
```{r}
# Create standard curve following kit instructions
standards <- tribble(
  ~std, ~BSA_ug.mL,
  "A",        2000,
  "B",        1500,
  "C",        1000,
  "D",         750,
  "E",         500,
  "F",         250,
  "G",         125,
  "H",          25,
  "I",           0
)

std_curve <- df %>%
  unnest(merged) %>%
  filter(grepl("Standard", colony_id)) %>%
  select(plate, Well, colony_id, abs562 = `562:562`) %>%
  rename(std = colony_id) %>%
  mutate(std = str_sub(std, 9, 9)) %>%
  #group_by(std) %>%
  #summarise(abs562 = mean(abs562)) %>%                       # calculate mean of standard duplicates
  #mutate(abs562.adj = abs562 - abs562[std == "I"]) %>%       # subtract blank absorbace value from all
  left_join(standards)

plot(std_curve$BSA_ug.mL, std_curve$abs562)
head(std_curve)

std_curve <- std_curve %>%
  filter(Well!="G8") 
```

Fit a linear model for each plate separately, calculate the coefficients, and then join them back together to match coefficients by plate number. We had one round bottom plate, so we need to treat this plate separately. 

Plate 1 (flat bottom)

```{r}
## Fit linear model for standard curve
std_curve_1<-std_curve%>%
  filter(plate=="20231207_Plate1")

 mod1 <- lm(BSA_ug.mL ~ abs562, data = std_curve_1)
# coef(mod)

## Fit nonlinear model for standard curve
mod1 <- nls(formula = BSA_ug.mL ~ z + a * exp(b * abs562), start = list(z = 0, a = 1, b = 1), data = std_curve_1)

fitted1 <- mod1 %>% broom::augment()

# Plot standard curve
std_curve_plot1 <- std_curve_1 %>%
  ggplot(aes(x = abs562, y = BSA_ug.mL)) +
  geom_point(color = "red", size = 3) 

std_curve_plot1 + 
  geom_line(data = fitted1, aes(x = abs562, y = .fitted)) +
  labs(title = "Standard curve - Plate 1")
```

Plate 2 and 3 (round bottom)

```{r}
## Fit linear model for standard curve
std_curve_23<-std_curve%>%
  filter(!plate=="20231207_Plate1")%>%
  filter(!c(plate=="20231207_Plate2" & Well=="H4"))%>%
  filter(!c(plate=="20231207_Plate3" & Well=="F3"))

#mod23 <- lm(BSA_ug.mL ~ abs562, data = std_curve_23)
# coef(mod)

## Fit nonlinear model for standard curve
mod23 <- nls(formula = BSA_ug.mL ~ z + a * exp(b * abs562), start = list(z = 0, a = 1, b = 1), data = std_curve_23)

fitted23 <- mod23 %>% broom::augment()

# Plot standard curve
std_curve_plot23 <- std_curve_23 %>%
  ggplot(aes(x = abs562, y = BSA_ug.mL)) +
  geom_point(color = "red", size = 3) 

std_curve_plot23 + 
  geom_line(data = fitted23, aes(x = abs562, y = .fitted)) +
  labs(title = "Standard curve - Plate 2 and 3")
```

# Calculate protein concentrations

We need to do this by plate specific standard curves. 

Plate 1 
```{r}
# Calculate protein concentration for all samples using standard curve
prot_plate1 <- df %>%
  unnest(merged) %>%
  filter(!grepl("Standard", colony_id)) %>%                     # Get just samples (not standards)
  filter(plate=="20231207_Plate1")%>%
  select(plate, Well, colony_id, abs562 = `562:562`) %>%        # Select only needed columns
  filter(!is.na(colony_id)) %>%                                 # Filter out empty wells
  filter(colony_id != "BK") %>%                                 # Filter out blank wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod1, newdata = data.frame(abs562 = .))))    # Use standard curve to convert absorbance to protein

std_curve_plot1 + 
  geom_point(data = prot_plate1, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "Plate 1 samples projected on standard curve")
```
POC-R11 is above the standard curve. Come back to this. 

Plate 2-3 
```{r}
# Calculate protein concentration for all samples using standard curve
prot_plate23 <- df %>%
  unnest(merged) %>%
  filter(!grepl("Standard", colony_id)) %>%                     # Get just samples (not standards)
  filter(!plate=="20231207_Plate1")%>%
  select(plate, Well, colony_id, abs562 = `562:562`) %>%        # Select only needed columns
  filter(!is.na(colony_id)) %>%                                 # Filter out empty wells
  filter(colony_id != "BK") %>%                                 # Filter out blank wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod23, newdata = data.frame(abs562 = .))))    # Use standard curve to convert absorbance to protein

std_curve_plot23 + 
  geom_point(data = prot_plate23, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "Plates 2 and 3 samples projected on standard curve")
```
POC-A6 is just above the standard curve. Come back to this. 

Merge data together 
```{r}
prot<-rbind(prot_plate1, prot_plate23)
```

Average technical replicates for each sample_id. 
```{r}
prot<-prot%>%
  group_by(colony_id, plate)%>%
  summarise(prot_ug.mL=mean(prot_ug.mL))
```

# Normalize to surface area
```{r}
# Surface area data
sa <- read.csv("output/physiology/surface_area/surface_area.csv")
colnames(sa)[1] <- "colony_id"

# Tissue homogenate volume data
homog_vols <- read_csv("data/physiology/homog_vols.csv") 
colnames(homog_vols)[1] <- "colony_id"

# Coral sample metadata
metadata <- read_csv("data/field_collection/collection_metadata.csv") 
colnames(metadata)[1] <- "colony_id"

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols, by="colony_id") %>%
  full_join(sa, by="colony_id")

# Join prot data with metadata
prot <- left_join(prot, metadata) %>%
  mutate(prot_ug = prot_ug.mL * total_slurry_ml,
         prot_ug.cm2 = prot_ug / surface.area.cm2,
         prot_mg.cm2 = prot_ug.cm2 / 1000)
```

# Plot results by species and stage

Plot observations. 
```{r}
# Plot all data points with mean ± se
prot %>%
  filter(!is.na(species)) %>%
  ggplot(aes(x = lifestage, y = prot_mg.cm2, color = species)) +
  facet_wrap(~species)  +
  coord_cartesian(ylim = c(0, 1.75))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean
```

Plot with labels for sample ID. 
```{r}
prot %>%
  filter(!is.na(species)) %>%
  ggplot(aes(x = lifestage, y = prot_mg.cm2, color = species)) +
  facet_wrap(~species)  +
  coord_cartesian(ylim = c(0, 1.75))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_jitter(width = 0.1) +  
  geom_text(aes(label=colony_id),hjust=0.1, vjust=0.1, size=2)
```
POC A6 was above the standard curve. It does not look like an outlier, although it is the lowest observation. 
POC-R11 was above the standard curve but is not an outlier from what I can tell. It is the highest observation. 

# Write data to output file
```{r}
# Write protein data to output file
prot <- prot %>%
  group_by(colony_id,species,lifestage) %>%
  summarise(prot_mg.cm2 = mean(prot_mg.cm2, na.rm = T)) %>%
  select(colony_id, species, lifestage, prot_mg.cm2) 

  write_csv(prot, file = "output/physiology/protein/protein.csv")
```

# Plotting and ANOVA analyses 

Attach temperature metadata. 
```{r}
sample.info <- read_csv(file = "data/p_r_rates/pr_sample_metadata.csv")
sample.info<-sample.info%>%rename(colony_id=sample_id)%>%select(colony_id, Temp.Cat)%>%filter(!grepl("BK", colony_id))

prot<-left_join(prot, sample.info)
```

View individual observations. 
```{r}
prot_plot1<-prot%>%
  
  ggplot(aes(x=Temp.Cat, y=prot_mg.cm2, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot1

ggsave("figures/physiology/protein/protein_dots.png", prot_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
prot_plot2<-prot%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(prot_mg.cm2, na.rm=TRUE), sd=sd(prot_mg.cm2, na.rm=TRUE), N=length(prot_mg.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot2

ggsave("figures/physiology/protein/protein_means.png", prot_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
prot_plot3<-prot%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(prot_mg.cm2, na.rm=TRUE), sd=sd(prot_mg.cm2, na.rm=TRUE), N=length(prot_mg.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot3

ggsave("figures/physiology/protein/protein_overall_means.png", prot_plot3, width = 6, height = 4)
```

## Run ANOVA model

```{r}
model1<-prot%>%
  
  aov(prot_mg.cm2~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Strong effects of species and lifestage, no interactions. Adult protein is higher than recruit. In POC and POR, adult protein is higher than recruit protein. No difference in ACR detected in posthoc tests. There is a difference in ACR-POR and POC-POR. No difference between ACR and POC at a species level. Between species, POC and ACR recruits are lower than POR recruits. No difference between ACR and POC recruits. ACR and POC adult are lower than POR adults. No difference in ACR and POC adults. 

