---
title: "Larval P & R Plotting and Analysis - Moorea 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pr_data<-read.csv("output/larvae_sdr/larvae_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pr_data<-pr_data[!is.na(pr_data$Type),]

#format columns
pr_data$Species<-as.factor(pr_data$Species)
pr_data$SDR<-as.factor(pr_data$SDR)
pr_data$Plate<-as.factor(pr_data$Plate)
pr_data$Temperature<-as.factor(pr_data$Temperature)
```

Look for outliers in the data.  

```{r}
boxplot(pr_data$P.nmol.org.min)
#pi_temp_data<-pi_temp_data%>%filter(P.nmol.org.min < 0.10)
#boxplot(pi_temp_data$P.nmol.org.min)

boxplot(pr_data$P.nmol.org.min~pr_data$PAR*pr_data$Species)
```

Set values at 0 for light runs if <0 indicating no photosynthesis occurred. Set values at 0 for dark runs if >0 indicating no respiration occurred. 
```{r}
pr_data<-pr_data%>%
  mutate(P.nmol.org.min=if_else(Species=="Pocillopora" & P.nmol.org.min<0 & PAR==150, 0,
                              if_else(Species=="Pocillopora" & P.nmol.org.min>0 & PAR==0, 0,
                                    if_else(Species=="Acropora" & P.nmol.org.min>0, 0, P.nmol.org.min))))

boxplot(pr_data$P.nmol.org.min~pr_data$PAR*pr_data$Species)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/larvae_sdr/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pr_data$Plate<-as.integer(pr_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pr_data<-left_join(pr_data,pr.temps)

#round to 0.1°C 
pr_data<-pr_data%>%
  mutate(Temp.C=round(Temp.C,1))

levels(as.factor(pr_data$Temp.C))
levels(as.factor(pr_data$Temperature))

pr.temps%>%group_by(Plate)%>%summarise(mean=mean(Temp.C))
```

Rename temperature categories. 
```{r}
pr_data<-pr_data%>%
  mutate(Temperature=if_else(Temperature=="27", "28.4", 
                           if_else(Temperature=="30", "29.4", 
                                   if_else(Temperature=="33", "31.5", NA))))

pr_data$Temperature
```

Reorder factor. 
```{r}
pr_data$PAR<-as.factor(pr_data$PAR)
pr_data$PAR
pr_data$Light<-as.factor(pr_data$Light)
pr_data$Light
```

# Calculate P gross and PR ratio metrics 

```{r}
pr_data<-pr_data%>%
  select(Temperature, Species, unique, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  rename(LEDR=`0`, Pnet=`150`)%>%
  mutate(LEDR=if_else(LEDR>0, 0, LEDR))%>% #if no oxygen consumption was observed, add 0 for respiration value
  mutate(LEDR=LEDR*-1)%>%
  mutate(Pgross=Pnet+LEDR)%>%
  mutate(Temperature=as.factor(Temperature))%>%
  mutate(Pgross=if_else(Pgross<0, 0, Pgross)) %>%
  mutate(PR_ratio=Pgross/LEDR)%>%
  mutate(PR_ratio=if_else(is.nan(PR_ratio), 0, PR_ratio))
```

# Calculate metrics normalized to average larval size 

For now, add a manual dummy value assuming Acropora are 10x larger than Pocillopora. 
```{r}
pr_data<-pr_data%>%
  mutate(size=if_else(Species=="Acropora", 1, 
                       if_else(Species=="Pocillopora", 0.1, NA)))%>% #normalize to size for POC and ACR dummy values
  mutate(LEDR_size=LEDR/size, Pnet_size=Pnet/size, Pgross_size=Pgross/size)
```

Prepare long data frame. 
```{r}
pr_long<-pr_data %>%
      mutate(PR_ratio=if_else(is.infinite(PR_ratio), NA, PR_ratio))%>%
    pivot_longer(names_to = "Metric", values_to = "Value", cols=c(Pnet, LEDR, Pgross, PR_ratio, Pnet_size, LEDR_size, Pgross_size))%>%
  select(Temperature, Species, unique, Metric, Value)
```

Remove well 3-B1 - respiration rates were weird in the light (but worked in the dark). So remove for now and we can come back and replace the light period with the dark period as this sample did not reach oxygen depletion. 
```{r}
pr_long<-pr_long%>%
  filter(!unique=="3 - B1")
```

Plot respiration rates for ACR only at 150 PAR. This is where respiration is taking place. By the time we get to the dark portion of the run oxygen is depleted and rates decrease. POC rates taken from 0 PAR portion of the run. 

# LEDR

Visualize means
```{r}
plots1<-pr_long%>%
    group_by(Species, Metric, Temperature)%>%
    filter(Metric %in% c("LEDR", "LEDR_size"))%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Temperature, y = mean, group=Temperature)) +
    facet_grid(~Species*Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Temperature, colour=Temperature), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab(expression(bold("Temperature (°C)"))) +  
    ylab(expression(bold(paste("LEDR (nmol ", O[2], " normalizer"^-1, "min"^-1, ")")))) +
    scale_colour_manual(values=c("blue3", "orange", "darkred"))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plots1

ggsave("figures/larvae_sdr/LEDR_means.png", plots1, dpi=300, w=7, h=4, units="in")
```

Run anova on individual normalized LEDR. 
```{r}
model<-pr_long%>%
  filter(Metric=="LEDR")%>%
  aov(Value~Species*Temperature, data=.)

summary(model)

emm<-emmeans(model, ~Temperature | Species)
pairs(emm)
```
Interactions of Species x Temperature on LEDR when individual normalized. 

No temp effects on individual normalized LEDR in POC. Strong decrease in above ambient temperatures in LEDR in ACR. 

Run anova on size normalized LEDR. 
```{r}
model<-pr_long%>%
  filter(Metric=="LEDR_size")%>%
  aov(Value~Species*Temperature, data=.)

summary(model)

emm<-emmeans(model, ~Temperature | Species)
pairs(emm)
```

Interaction of species x temperature on size normalized data. But no main effect of species - they are similar rates when normalized to size. 

No temperature differences in size normalized POC rates. Maintained decrease in LEDR in ACR normalized to size. 

# Pgross (POC only)

Visualize means
```{r}
plots2<-pr_long%>%
    group_by(Species, Metric, Temperature)%>%
    filter(Metric %in% c("Pgross", "Pgross_size"))%>%
    filter(Species=="Pocillopora")%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Temperature, y = mean, group=Temperature)) +
    facet_grid(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Temperature, colour=Temperature), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab(expression(bold("Temperature (°C)"))) +  
    ylab(expression(bold(paste(P[Gross], " (nmol ", O[2], " normalizer"^-1, "min"^-1, ")")))) +
    scale_colour_manual(values=c("blue3", "orange", "darkred"))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plots2

ggsave("figures/larvae_sdr/Pgross_means.png", plots2, dpi=300, w=4, h=4, units="in")
```

Run anova on individual normalized Pgross.  
```{r}
model<-pr_long%>%
  filter(Species=="Pocillopora")%>%
  filter(Metric=="Pgross")%>%
  aov(Value~Temperature, data=.)

summary(model)
```
No effect of temp on P gross in POC. 

Run anova on size normalized P gross.  
```{r}
model<-pr_long%>%
  filter(Metric=="Pgross_size")%>%
  aov(Value~Temperature, data=.)

summary(model)
```
No temperature differences in size normalized POC rates.

No effects seem to be due to POC variation at ambient - look into this. Difference between 29.4 and 31.5 are very strong. 

# Pnet (POC only)

Visualize means
```{r}
plots3<-pr_long%>%
    group_by(Species, Metric, Temperature)%>%
    filter(Metric %in% c("Pnet", "Pnet_size"))%>%
    filter(Species=="Pocillopora")%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Temperature, y = mean, group=Temperature)) +
    facet_grid(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Temperature, colour=Temperature), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab(expression(bold("Temperature (°C)"))) +  
    ylab(expression(bold(paste(P[Net], " (nmol ", O[2], " normalizer"^-1, "min"^-1, ")")))) +
    scale_colour_manual(values=c("blue3", "orange", "darkred"))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plots3

ggsave("figures/larvae_sdr/Pnet_means.png", plots3, dpi=300, w=4, h=4, units="in")
```

Run anova on individual normalized Pnet.  
```{r}
model<-pr_long%>%
  filter(Species=="Pocillopora")%>%
  filter(Metric=="Pnet")%>%
  aov(Value~Temperature, data=.)

summary(model)

emm<-emmeans(model, ~Temperature)
pairs(emm)
```
Effect of temp on P net. Lower at 31.5. 

Run anova on size normalized P net  
```{r}
model<-pr_long%>%
  filter(Metric=="Pnet_size")%>%
  aov(Value~Temperature, data=.)

summary(model)
```
No temperature differences in size normalized P net POC rates.

# P:R ratio (POC only)

Visualize means
```{r}
plots4<-pr_long%>%
    group_by(Species, Metric, Temperature)%>%
    filter(Metric %in% c("PR_ratio"))%>%
    filter(Species=="Pocillopora")%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Temperature, y = mean, group=Temperature)) +
    facet_grid(~Metric)+
    geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Temperature, colour=Temperature), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab(expression(bold("Temperature (°C)"))) +  
    ylab(expression(bold(paste("P:R Ratio")))) +
    scale_colour_manual(values=c("blue3", "orange", "darkred"))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plots4

ggsave("figures/larvae_sdr/PR_ratio_means.png", plots4, dpi=300, w=4, h=4, units="in")
```

Run anova on PR ratios.  
```{r}
model<-pr_long%>%
  filter(Species=="Pocillopora")%>%
  filter(Metric=="PR_ratio")%>%
  aov(Value~Temperature, data=.)

summary(model)

emm<-emmeans(model, ~Temperature)
pairs(emm)
```
Effect of temp on PR ratios. Lower at 31.5. 