---
title: "Symbiont density analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")

# load packages
library(tidyverse)
library(emmeans)
```

# Import data
```{r}
# Cell count data
sym_counts <- read_csv("data/physiology/cell_density/cells.csv")

# Surface area data
sa <- read.csv("output/physiology/surface_area/surface_area.csv") 

# Tissue homogenate volume data
homog_vols <- read_csv("data/physiology/homog_vols.csv") %>% select(sample_id, total_slurry_ml, sym_vol_ml, host_vol_ml, resuspended_sym_vol_ml)

# Coral sample metadata
metadata <- read_csv("data/p_r_rates/pr_sample_metadata.csv") %>% select(sample_id, Species, Lifestage, Temp.Cat) %>% filter(!str_detect(sample_id, "^BK"))

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols) %>%
  full_join(sa) %>% filter(!is.na(Species))
```


# Calculate cells per square centimeter
```{r}
# Calculate mean counts for each sample
sym_counts <- sym_counts %>%
  select(sample_id, squares_counted, matches("count[0-6]")) %>%
  gather("rep", "count", -sample_id, -squares_counted) %>%
  group_by(sample_id, squares_counted) %>%
  summarise(mean_count = mean(count, na.rm = TRUE))

# Join mean counts with sample metadata
sym_counts <- full_join(sym_counts, metadata)


# homogenate volume = total tissue
# sym aliquot volume = total tissue from which cells were taken
# sym resuspended volume = dilution of cells from total pellet 

#count * 10000 / squares counted = cells per ml
#cells per ml * resuspended volume = total cells in aliquot
#total cells in slurry = total cells * (total slurry / aliquot volume)
# cells per unit sa = total cells in slurry / sa 

# Normalize counts by homogenate volume and surface area

sym_counts <- sym_counts %>%
  mutate(cells.ml = mean_count * 10000 / squares_counted,
         aliquot.cells = cells.ml * resuspended_sym_vol_ml, 
         total.cells = aliquot.cells * (total_slurry_ml/sym_vol_ml),
         cells.cm2 = total.cells / surface.area.cm2)
```

# Plot data
```{r}
sym_counts %>%

  ggplot(aes(x = Lifestage, y = cells.cm2 / 10^6), colour=as.factor(Temp.Cat)) +
  labs(x = "",y = "Cell Density (Cells/cm2 * 10^6)") +
  facet_wrap(~ Species) +
  geom_jitter(aes(colour=as.factor(Temp.Cat)),width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.1) +
  stat_summary(fun = mean, geom = "point", color = "black")+
  scale_colour_manual(values=c("blue2", "orange", "darkred"))+
  theme_classic()# Plot mean
```

```{r}
sample_plot<-sym_counts %>%

  ggplot(aes(x = Lifestage, y = cells.cm2 / 10^6), colour=Lifestage) +
  labs(x = "",y = "Cell Density (Cells/cm2 * 10^6)") +
  facet_wrap(~ Species) +
  geom_jitter(aes(colour=Lifestage),width = 0.1) +                                            # Plot all points
  theme_classic() + 
  geom_text(aes(label=sample_id),hjust=0, vjust=0, size=2);sample_plot

ggsave("figures/physiology/cell_density/cells_samples.png", sample_plot, dpi=300, w=6, h=4, units="in")
```

Generate final figures 

Plot with mean and standard error by lifestage and species. 
```{r}
plot1<-sym_counts %>%
    filter(!is.na(cells.cm2))%>%
    group_by(Species, Lifestage)%>%
    dplyr::summarise(mean=mean(cells.cm2, na.rm=TRUE), sd=sd(cells.cm2, na.rm=TRUE), N=length(cells.cm2), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Lifestage, y = mean / 10^6, group=Lifestage)) +
    facet_grid(~Species)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Lifestage), size=6, position = position_dodge(0.4), color="darkgray") + 
    geom_errorbar(aes(ymin=(mean/ 10^6)-(se/ 10^6), ymax=(mean/ 10^6)+(se/ 10^6), group=Lifestage), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Lifestage") + 
    ylab(expression(bold(paste("Cell Density (10"^6, "cells cm"^-2, ")")))) +
    ggtitle("")+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plot1

ggsave("figures/physiology/cell_density/cells_means.png", plot1, dpi=300, w=6, h=4, units="in")
```

Plot by individual observation with colour by temperature exposure. 
```{r}
plot2<-sym_counts %>%
    filter(!is.na(cells.cm2))%>%
    mutate(Temp.Cat=as.factor(Temp.Cat))%>%
    
    ggplot(., aes(x = Lifestage, y = cells.cm2 / 10^6, group=Lifestage, color=Temp.Cat)) +
    facet_grid(~Species)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Lifestage), size=3, position = position_jitterdodge(0.5)) + 
    xlab("Lifestage") + 
    ylab(expression(bold(paste("Cell Density (10"^6, "cells cm"^-2, ")")))) +
    scale_color_manual(values=c("blue2", "orange", "red3"))+
    ggtitle("")+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plot2

ggsave("figures/physiology/cell_density/cells_dots.png", plot2, dpi=300, w=6, h=4, units="in")
```

# Run ANOVA tests 

```{r}
model1<-sym_counts%>%
  
  aov(cells.cm2~Temp.Cat*Species*Lifestage, data=.)

summary(model1)
```

No effect of temp, run only with species and lifesetage. 
```{r}
model2<-sym_counts%>%
  
  aov(cells.cm2~Species*Lifestage, data=.)

summary(model2)

emm<-emmeans(model2, ~Lifestage | Species)
pairs(emm)
```

There is an effect of species and lifestage.  

Recruits have lower cell density than adults in POC and POR species. No difference in Acropora. 

# Plot recruit symbiont density as a function of size. 

```{r}
plot3<-sym_counts %>%
    filter(!is.na(cells.cm2))%>%
    mutate(Temp.Cat=as.factor(Temp.Cat))%>%
    filter(Lifestage=="Recruit")%>%
    
    ggplot(., aes(x = surface.area.cm2, y = cells.cm2 / 10^6, group=Lifestage)) +
    facet_grid(~Species, scales="free_x")+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=Lifestage), size=3) + 
    xlab("Surface Area") + 
    ylab(expression(bold(paste("Cell Density (10"^6, "cells cm"^-2, ")")))) +
    #scale_color_manual(values=c("blue2", "orange", "red3"))+
    geom_smooth(method="lm", colour="darkgray", se=FALSE, linewidth=1)+
    ggtitle("")+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      title = element_text(size=14, face="bold")
      ); plot3

ggsave("figures/physiology/cell_density/cells_size.png", plot3, dpi=300, w=6, h=4, units="in")
```

Run analysis 
```{r}
model3<-sym_counts %>%
    filter(!is.na(cells.cm2))%>%
    mutate(Temp.Cat=as.factor(Temp.Cat))%>%
    filter(Lifestage=="Recruit")%>%
  
    aov(cells.cm2~surface.area.cm2*Species, data=.)

summary(model3)
```

Significant relationship between cell density and recruit size for each species. Difference in relationship between size and density between species. 

Visually there is a strong positive relationship between recruit size and cell density per unit size in POC and POR, but not ACR. 

# Output files

Output data to file.  

```{r}
sym_counts %>%
  select(sample_id, Lifestage, Species, Temp.Cat, cells.cm2) %>%
  write_csv(file = "output/physiology/cell_density/symbiont_densities.csv")

```

# View symbiont cell density normalized to total protein 

Read in cell density and total protein and merge. 
```{r}
prot<-read_csv(file="output/physiology/protein/protein.csv")%>%select(colony_id, prot_mg.cm2)
colnames(prot)[1] <- "sample_id"

sym_counts<-left_join(sym_counts, prot)
```

Calculate total chl normalized to cells and normalized to protein. 
```{r}
sym_counts<-sym_counts%>%
  mutate(cells.mg.prot=cells.cm2/prot_mg.cm2)
```

View individual observations. 
```{r}
cells_prot_plot1<-sym_counts%>%
  
  ggplot(aes(x=Temp.Cat, y=cells.mg.prot, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
    ylab(expression(bold(paste("Cell Density (10"^6, "cells mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); cells_prot_plot1

ggsave("figures/physiology/cell_density/cells_protein_dots.png", cells_prot_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
cells_prot_plot2<-sym_counts%>%
  group_by(Species, Lifestage, Temp.Cat)%>%
  mutate(mean=mean(cells.mg.prot, na.rm=TRUE), sd=sd(cells.mg.prot, na.rm=TRUE), N=length(cells.mg.prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
    ylab(expression(bold(paste("Cell Density (10"^6, "cells mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); cells_prot_plot2

ggsave("figures/physiology/cell_density/cells_protein_means.png", cells_prot_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
cells_prot_plot3<-sym_counts%>%
  group_by(Species, Lifestage)%>%
  mutate(mean=mean(cells.mg.prot, na.rm=TRUE), sd=sd(cells.mg.prot, na.rm=TRUE), N=length(cells.mg.prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Lifestage, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
    ylab(expression(bold(paste("Cell Density (10"^6, "cells mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); cells_prot_plot3

ggsave("figures/physiology/cell_density/cells_protein_overall_means.png", cells_prot_plot3, width = 6, height = 4)
```

## Run ANOVA models

```{r}
model1<-sym_counts%>%
  
  aov(cells.mg.prot~Species*Lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ Species)
pairs(emm)

emm<-emmeans(model1, ~ Species*Lifestage)
pairs(emm)
```
Species effects. ACR values are higher than other species.  
