---
title: "Oxygen flux rate extractions "
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("LoLinR")) install_github('colin-olito/LoLinR') 

## load libraries
library(devtools)
library(LoLinR)
library(tidyverse)
library(lubridate)
library(cowplot)
library(broom)
library(plotrix)
library(Hmisc)
library(rTPC)
library(nls.multstart)
library(emmeans)

## libraries for parallel processing
library(future)
library(furrr)
```

## Import metadata
```{r, warning = FALSE}
path.p <- "data/p_r_rates/runs" #the location of all your respirometry files 

# List data files
file.names <- list.files(path = path.p, pattern = "csv$")  # list all csv file names in the folder
file.names <- file.names[!grepl("metadata", file.names)]   # omit metadata from files to be read in as data

# Load PI curve sample metadata (i.e., which corals were in which runs)
sample.info <- read_csv(file = "data/p_r_rates/pr_sample_metadata.csv")

# Load PI curve run metadata (i.e., light levels and interval times for each run)
run.info <- read_csv(file = "data/p_r_rates/pr_run_metadata_test.csv")

# Join all coral and run metadata
metadata <- full_join(sample.info, run.info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = "Tahiti"))
metadata <- metadata %>% filter(!Light_Level==3)

# Select only certain columns
metadata <- metadata %>%
  select(Species, Lifestage, sample_id, Run, Temp.Cat, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Value, Light_Level)
```


# Read in all data files
```{r, results="hide", message=FALSE}
#identify the sample id from the file name and read in data
df <- tibble(file.name = file.names) %>%
  mutate(sample_id = gsub("_.*", "", file.name),                              # Get sample_id from filename
          info = map(sample_id, ~filter(metadata, sample_id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip=1, col_types = cols(.default = "d", Time = "t"), col_select = c(Time, Value, Temp))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}

unnest<-unnest_legacy
nest<-nest_legacy

#Use start time of each light step from the metadata to separate data by light stop
df <- df %>%
  dplyr::mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Value)))})) %>%
  dplyr::mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Value")))

```

### Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))

# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample_id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Value), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample_id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Value), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("         Example plot: all data", "      Example plot: thinned data"))
```

#### The full or thinned data plot for any sample can be accessed like this:
```
df %>%
  filter(sample_id == "Sample-Name-Here") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r} 
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.4, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
future::plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Value) %>%
    do(rankLcRg = fit_reg(.))
  }))

#NEED TO ADD IN A LINE TO EXTRACT MEAN TEMPERATURE OF EACH LIGHT LEVEL
# df <- df %>%
#   mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
#     group_by(., Light_Value) %>%
#     do(rankLcRg = fit_reg(.), 
#        MeanTemp = summarise(., mean(Temp))
#   }))

## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample_id, interval_number) {
  df %>%
    filter(sample_id == sample_id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```

#### The diagnostics for any regression can be plotted by specifying a sample_id and the number of the light curve interval:
```{r}
# Define function to pull out and plot regression diagnostics

#THIS FUNCTION IS NOT WORKING BY COLONYID
# pdf("output/Past-D1_test.pdf")
plot_rankLcRg("ACR-A10", 1)
plot_rankLcRg("ACR-A10", 2)
plot_rankLcRg("POR-A12", 1)
plot_rankLcRg("POR-A12", 2)
# plot_rankLcRg("Past-D1", 3)
# plot_rankLcRg("Past-D1", 4)
# plot_rankLcRg("Past-D1", 5)
# plot_rankLcRg("Past-D1", 6)
# plot_rankLcRg("Past-D1", 7)
# plot_rankLcRg("Past-D1", 8)
# plot_rankLcRg("Past-D1", 9)
# plot_rankLcRg("Past-D1", 10)
# dev.off()
```


DO THIS NEXT 

Generate and export a diagnostic plot for each temperature and light level. 
```{r}

```



### Extract slope of best regression for each interval for each sample
```{r}
#extract slope as rate
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))

#verify sample numbers
unique(df.out[[2]])
length(unique(df.out[[2]]))

#select only the essential columns
xx <- select(df.out,sample_id, Light_Value, micromol.L.s)

#add a grouping id for each colony at each light level
xx$grouping.id <- paste0(xx$sample_id, "-",xx$Light_Value)
nrow(xx)
nrow(distinct(xx))

#Check to see if all samples have the expected number of light steps
unique(xx[[4]])
xxx <- xx %>%
  group_by(sample_id) %>%
  summarise(light_steps = n())

#select only the essential columns
mx <- select(metadata, Species, Lifestage, sample_id, Run, Chamber.Vol.L, Temp.Cat)
nrow(mx)
nrow(distinct(df))

#join rates with metadata
pr <- left_join(xx, mx, by="sample_id")
length(unique(pr$sample_id))
nrow(distinct(pr))
pr <-distinct(pr)
  
# Write raw data to output file
write.csv(pr, "output/p_r_rates/pr_curve_extracted_rates_Raw.csv")
```


# Adjust rates by chamber volume, subtract blank, and normalize to surface area
```{r}
# Correct for chamber volume 
pr <- pr %>% mutate(micromol.s = micromol.L.s * Chamber.Vol.L)
length(unique(pr$sample_id))

# Correct for blank rates
# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BK", sample_id)) %>%
  group_by(Temp.Cat, Light_Value) %>%
  summarise(micromol.s.blank=mean(micromol.s))%>%
  mutate(blank_id=paste0(Temp.Cat,"-",Light_Value))

### generate a key for the blank id
pr <- pr %>%
  mutate(blank_id=paste0(Temp.Cat,"-",Light_Value))
length(unique(pr$sample_id))

#plot blank values
blanks %>% ggplot(aes(x=as.numeric(Light_Value), y=micromol.s.blank,colour = as.factor(Temp.Cat)))+
  geom_point()+
  geom_line()

#examine the effects of light and temp on the blank rates
anova(lm(micromol.s.blank~as.factor(Light_Value)*Temp.Cat, data=blanks))

#join the data and the mean of the blanks per temperature for each specific light level
pr  <- left_join(pr ,blanks,by = "blank_id")
length(unique(pr$sample_id))

#subtract temp and light specific blank values from samples
pr <- pr %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BK", sample_id))
length(unique(pr$sample_id))

# Import surface area data
sa <- read.csv("output/physiology/surface_area/surface_area.csv")

# Join surface area with rest of data
pr <- left_join(pr, select(sa, sample_id, surface.area.cm2))
length(unique(pr$sample_id))

pr  <- dplyr::inner_join(
   pr ,
   dplyr::select(sa, -any_of(names(pr)), sample_id),
   by = "sample_id"
)
length(unique(pr$sample_id))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)
length(unique(pr$sample_id))
```

# Write extracted rates to output file
```{r}
# Select variables to write to file
pr.out <- pr %>% 
  select(Species, Lifestage, sample_id, Light_Value.x, Temp.Cat.x, Run, micromol.cm2.s, micromol.cm2.h)
length(unique(pr.out$sample_id))

# Write to output file
write.csv(pr.out, "output/p_r_rates/pr_curve_extracted_rates.csv")
```

Add in surface area into data frame 
```{r}
pr.out$surface.area<-sa$surface.area.cm2[match(pr.out$sample_id, sa$sample_id)]
```

Visualize LEDR
```{r}
#plot  values
LEDR <- pr.out %>% filter(Light_Value.x==0)
  
ledr_plot<-LEDR %>% ggplot(aes(x=as.factor(Temp.Cat.x), y=-1*micromol.cm2.h, colour = Lifestage))+
  geom_point(position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  labs(x = "Temp 째C")+
  labs(y = "Rdark (micromol.cm2.h)")+
  theme_classic();ledr_plot

ggsave("figures/p_r_rates/Rdark.pdf", ledr_plot, width = 8, height = 4)

# view by means 

ledr_plot_2<-LEDR %>% 
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  summarise(mean=mean(micromol.cm2.h, na.rm=TRUE), se=sd(micromol.cm2.h, na.rm=TRUE)/sqrt(length(micromol.cm2.h)))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=-1*mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.5), size=2)+
  geom_errorbar(aes(ymin=(-1*mean)-se, ymax=(-1*mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.5))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 2)+
  labs(x = "Temp 째C")+
  labs(y = "LEDR (micromol.cm2.h)")+
  theme_classic();ledr_plot_2

ggsave("figures/p_r_rates/Rdark_means.pdf", ledr_plot_2, width = 8, height = 4)
```


Visualize Pmax
```{r}
#plot  values
Pmax <- pr.out %>% filter(Light_Value.x==550)
  
pmax_plot<-Pmax %>% ggplot(aes(x=as.factor(Temp.Cat.x), y=micromol.cm2.h, colour = Lifestage))+
  geom_point(aes(size=surface.area), position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  #geom_text(aes(label=sample_id),hjust=0, vjust=0)+
  labs(x = "Temp 째C")+
  labs(y = "Pmax (micromol.cm2.h)")+
  theme_classic(); pmax_plot

ggsave("figures/p_r_rates/Pmax.pdf", pmax_plot, width = 8, height = 4)

# view by means 

pmax_plot_2<-Pmax %>% 
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  summarise(mean=mean(micromol.cm2.h, na.rm=TRUE), se=sd(micromol.cm2.h, na.rm=TRUE)/sqrt(length(micromol.cm2.h)))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.5), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.5))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  #ylim(0, 2)+
  labs(x = "Temp 째C")+
  labs(y = "Pmax (micromol.cm2.h)")+
  theme_classic();pmax_plot_2

ggsave("figures/p_r_rates/Pmax_means.pdf", pmax_plot_2, width = 8, height = 4)

```

View by size. 

```{r}
#plot  pmax 
Pmax_recruit <- pr.out %>% filter(Light_Value.x==550) %>% filter(Lifestage=="Recruit") %>% filter(surface.area<10)
  
pmax_plot_recruit<-Pmax %>% ggplot(aes(x=surface.area, y=micromol.cm2.h, colour = as.factor(Temp.Cat.x)))+
  geom_point(position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Temp.Cat.x), se=FALSE)+

  labs(x = "Surface Area (cm2)")+
  labs(y = "Pmax (micromol.cm2.h)")+
  theme_classic(); pmax_plot_recruit

ggsave("figures/p_r_rates/Pmax_recruits.pdf", pmax_plot_recruit, width = 8, height = 4)
```

Run ANOVA

Pmax
```{r}
model1<-aov(micromol.cm2.h~Lifestage*Species*Temp.Cat.x, data=Pmax)
summary(model1)

emm<-emmeans(model1, ~Lifestage | Species)
pairs(emm)
```

LEDR
```{r}
model2<-aov(micromol.cm2.h~Lifestage*Species*Temp.Cat.x, data=LEDR)
summary(model2)

emm<-emmeans(model2, ~Lifestage | Species)
pairs(emm)
```
