---
title: "Oxygen flux rate extractions "
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("LoLinR")) install_github('colin-olito/LoLinR') 

## load libraries
library(devtools)
library(LoLinR)
library(tidyverse)
library(lubridate)
library(cowplot)
library(broom)
library(plotrix)
library(Hmisc)
library(rTPC)
library(nls.multstart)
library(emmeans)

## libraries for parallel processing
library(future)
library(furrr)
```

## Import data
```{r, warning = FALSE}
pr<-read_csv(file="output/p_r_rates/pr_curve_extracted_rates.csv")
sa<-read_csv(file="output/physiology/surface_area/surface_area.csv")
```

Add in surface area into pr data frame 
```{r}
pr$surface.area<-sa$surface.area.cm2[match(pr$sample_id, sa$sample_id)]
```

# View by treatment 

Visualize LEDR
```{r}
#plot  values
LEDR <- pr %>% filter(Light_Value.x==0)
  
ledr_plot<-LEDR %>% ggplot(aes(x=as.factor(Temp.Cat.x), y=-1*micromol.cm2.h, colour = Lifestage))+
  geom_point(position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  labs(x = "Temp 째C")+
  labs(y = "Rdark (micromol.cm2.h)")+
  theme_classic();ledr_plot

ggsave("figures/p_r_rates/Rdark.png", ledr_plot, width = 8, height = 4)

# view by means 

ledr_plot_2<-LEDR %>% 
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  summarise(mean=mean(micromol.cm2.h, na.rm=TRUE), se=sd(micromol.cm2.h, na.rm=TRUE)/sqrt(length(micromol.cm2.h)))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=-1*mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.5), size=2)+
  geom_errorbar(aes(ymin=(-1*mean)-se, ymax=(-1*mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.5))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 2)+
  labs(x = "Temp 째C")+
  labs(y = "LEDR (micromol.cm2.h)")+
  theme_classic();ledr_plot_2

ggsave("figures/p_r_rates/Rdark_means.png", ledr_plot_2, width = 8, height = 4)
```

Visualize Pmax
```{r}
#plot  values
Pmax <- pr %>% filter(Light_Value.x==550)
  
pmax_plot<-Pmax %>% ggplot(aes(x=as.factor(Temp.Cat.x), y=micromol.cm2.h, colour = Lifestage))+
  geom_point(aes(size=surface.area), position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  #geom_text(aes(label=sample_id),hjust=0, vjust=0)+
  labs(x = "Temp 째C")+
  labs(y = "Pmax (micromol.cm2.h)")+
  theme_classic(); pmax_plot

ggsave("figures/p_r_rates/Pmax.png", pmax_plot, width = 8, height = 4)

# view by means 

pmax_plot_2<-Pmax %>% 
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  summarise(mean=mean(micromol.cm2.h, na.rm=TRUE), se=sd(micromol.cm2.h, na.rm=TRUE)/sqrt(length(micromol.cm2.h)))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.5), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.5))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1)+
  #ylim(0, 2)+
  labs(x = "Temp 째C")+
  labs(y = "Pmax (micromol.cm2.h)")+
  theme_classic();pmax_plot_2

ggsave("figures/p_r_rates/Pmax_means.png", pmax_plot_2, width = 8, height = 4)

```

# View by recruit size 

View by size. 

```{r}
#plot  pmax 
Pmax_recruit <- pr %>% filter(Light_Value.x==550) %>% filter(Lifestage=="Recruit") %>% filter(surface.area<10)
  
pmax_plot_recruit<-Pmax %>% ggplot(aes(x=surface.area, y=micromol.cm2.h, colour = as.factor(Temp.Cat.x)))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Temp.Cat.x), se=FALSE)+
  geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  labs(x = "Surface Area (cm2)")+
  labs(y = "P (micromol.cm2.h)")+
  theme_classic()+
  theme(legend.position="none"); pmax_plot_recruit

ggsave("figures/p_r_rates/Pmax_recruits.png", pmax_plot_recruit, width = 6, height = 4)
```

# Statistical tests 

Run ANOVA

Pmax
```{r}
model1<-aov(micromol.cm2.h~Lifestage*Species*Temp.Cat.x, data=Pmax)
summary(model1)

emm<-emmeans(model1, ~Lifestage | Species)
pairs(emm)
```

LEDR
```{r}
model2<-aov(micromol.cm2.h~Lifestage*Species*Temp.Cat.x, data=LEDR)
summary(model2)

emm<-emmeans(model2, ~Lifestage | Species)
pairs(emm)
```

There is a lifestage x species effect for both LEDR and P. 

# Run correlations between P and R and normalize to cell density 

Read in data
```{r}
sym<-read_csv(file="output/physiology/cell_density/symbiont_densities.csv")
```

Merge data and normalize rates to per cell. 
```{r}
pr$cells.cm2<-sym$cells.cm2[match(pr$sample_id, sym$sample_id)]
hist(pr$cells.cm2)

pr_wide<-pr%>%
  select(Species, Lifestage, sample_id, Temp.Cat.x, micromol.cm2.h, surface.area, cells.cm2, Light_Value.x)%>%
  pivot_wider(names_from=Light_Value.x, values_from=micromol.cm2.h)%>%
  rename(LEDR=`0`, P=`550`)%>%
  mutate(LEDR=LEDR*-1)%>%
  mutate(LEDR_cell=LEDR/cells.cm2)%>%
  mutate(P_cell=P/cells.cm2)
```

Plot correlation between P and R
```{r}
corr1<-pr_wide%>%
  
  ggplot(aes(x=LEDR, y=P, colour = as.factor(Temp.Cat.x)))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  #geom_smooth(method="lm", aes(group=Temp.Cat.x), se=FALSE)+
  geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  labs(x = "LEDR (micromol.cm2.h)")+
  labs(y = "P (micromol.cm2.h)")+
  theme_classic()+
  theme(legend.position="none"); corr1

ggsave("figures/p_r_rates/P_LEDR_correlation.png", corr1, width = 6, height = 4)
```

There is a strong correlation between LEDR and P. 
```{r}
model3<-lm(P~LEDR*Species, data=pr_wide)
summary(model3)
```
The relationship between P and LEDR is strong but not different between species. 

# View P and LEDR normalized to symbiont cell density 

## P per cell 

Plot P per cell. Identify outliers 
```{r}
hist(pr_wide$P_cell)
```

View individual observations. 
```{r}
cell_plot1<-pr_wide%>%
  filter(P_cell<0.00002)%>%
  
  ggplot(aes(x=Temp.Cat.x, y=P_cell, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  labs(x = "Lifestage")+
  labs(y = "P (micromol.cell.h)")+
  theme_classic()+
  theme(legend.position="none"); cell_plot1

ggsave("figures/p_r_rates/P_cell_dots.png", cell_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
cell_plot2<-pr_wide%>%
  filter(P_cell<0.00002)%>%
  filter(!is.na(P_cell))%>%
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  mutate(mean=mean(P_cell, na.rm=TRUE), sd=sd(P_cell, na.rm=TRUE), N=length(P_cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.5))+
    geom_point(position=position_dodge(0.5), alpha=1, size=3)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="black", width=0, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  labs(x = "Temperature")+
  labs(y = "P (micromol.cell.h)")+
  theme_classic()+
  theme(legend.position="none"); cell_plot2

ggsave("figures/p_r_rates/P_cell_means.png", cell_plot2, width = 6, height = 4)
```

Analyze P per cell 

Run anova. 
```{r}
model4<-pr_wide%>%
  filter(P_cell<0.00002)%>%
  
  aov(P_cell~Species*Lifestage*Temp.Cat.x, data=.)

summary(model4)

emm<-emmeans(model4, ~Lifestage | Species)
pairs(emm)
```
Effect of species and lifestage on P normalized to per cell rates. Recruits look higher than adults. Particularly in POC. No strong temperature effects. 

## LEDR per cell 

Plot LEDR per cell 

View individual observations. 
```{r}
cell_plot3<-pr_wide%>%
  filter(LEDR_cell<0.000025)%>%
  filter(LEDR_cell>0)%>%
  
  ggplot(aes(x=Temp.Cat.x, y=LEDR_cell, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  labs(x = "Lifestage")+
  labs(y = "LEDR (micromol.cell.h)")+
  theme_classic()+
  theme(legend.position="none"); cell_plot3

ggsave("figures/p_r_rates/LEDR_cell_dots.png", cell_plot3, width = 6, height = 4)
```

View mean observations. 
```{r}
cell_plot4<-pr_wide%>%
  filter(LEDR_cell<0.000025)%>%
  filter(LEDR_cell>0)%>%
  filter(!is.na(LEDR_cell))%>%
  group_by(Species, Lifestage, Temp.Cat.x)%>%
  mutate(mean=mean(LEDR_cell, na.rm=TRUE), sd=sd(LEDR_cell, na.rm=TRUE), N=length(LEDR_cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=as.factor(Temp.Cat.x), y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.5))+
    geom_point(position=position_dodge(0.5), alpha=1, size=3)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="black", width=0, position=position_dodge(0.5))+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  labs(x = "Temperature")+
  labs(y = "LEDR (micromol.cell.h)")+
  theme_classic()+
  theme(legend.position="none"); cell_plot4

ggsave("figures/p_r_rates/LEDR_cell_means.png", cell_plot4, width = 6, height = 4)
```

Analyze R per cell 
Run anova. 
```{r}
model5<-pr_wide%>%
  filter(LEDR_cell<0.000025)%>%
  filter(LEDR_cell>0)%>%
  
  aov(LEDR_cell~Species*Lifestage*Temp.Cat.x, data=.)

summary(model5)
```
No difference in LEDR normalized to cells. 


