---
title: "Oxygen flux rate extractions "
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("LoLinR")) install_github('colin-olito/LoLinR') 

## load libraries
library(devtools)
library(LoLinR)
library(tidyverse)
library(lubridate)
library(cowplot)
library(broom)
library(plotrix)
library(Hmisc)
library(rTPC)
library(nls.multstart)
library(emmeans)

## libraries for parallel processing
library(future)
library(furrr)
```

# Import data
```{r, warning = FALSE}
pr<-read_csv(file="output/p_r_rates/pr_curve_extracted_rates.csv")
sa<-read_csv(file="output/physiology/surface_area/surface_area.csv")
prot<-read_csv(file="output/physiology/protein/protein.csv")%>%rename(sample_id=colony_id)
```

Add in surface area and protein into pr data frame 
```{r}
pr$surface.area<-sa$surface.area.cm2[match(pr$sample_id, sa$sample_id)]

pr$prot_mg.cm2<-prot$prot_mg.cm2[match(pr$sample_id, prot$sample_id)]
hist(pr$prot_mg.cm2)
```

Load in cell density to normalize to cell density. 

```{r}
sym<-read_csv(file="output/physiology/cell_density/symbiont_densities.csv")
```

Merge data and normalize rates to per cell and 
```{r}
pr$cells.cm2<-sym$cells.cm2[match(pr$sample_id, sym$sample_id)]
hist(pr$cells.cm2)
```

# Calculate P gross and PR ratio and normalize to cells and protein

Calculate P gross as P net - LEDR and PR ratio as Pgross/LEDR. 
```{r}
pr_wide<-pr%>%
  select(Species, Lifestage, sample_id, Temp.Cat.x, micromol.cm2.h, surface.area, cells.cm2, prot_mg.cm2, Light_Value.x)%>%
  pivot_wider(names_from=Light_Value.x, values_from=micromol.cm2.h)%>%
  rename(LEDR=`0`, Pnet=`550`)%>%
  mutate(LEDR=if_else(LEDR>0, 0, LEDR))%>% #if no oxygen consumption was observed, add 0 for respiration value
  mutate(LEDR=LEDR*-1)%>%
  mutate(Pgross=Pnet+LEDR)%>%
  mutate(LEDR_cell=LEDR/cells.cm2)%>%
  mutate(Pnet_cell=Pnet/cells.cm2)%>%
  mutate(Pgross_cell=Pgross/cells.cm2)%>%
  mutate(LEDR_prot=LEDR/prot_mg.cm2)%>%
  mutate(Pnet_prot=Pnet/prot_mg.cm2)%>%
  mutate(Pgross_prot=Pgross/prot_mg.cm2)%>%
  rename(Temperature=Temp.Cat.x)%>%
  mutate(Temperature=as.factor(Temperature))%>%
  mutate(PR_ratio=Pgross/LEDR)
```

# View LEDR, Pnet, and Pgross per cm2 by treatment 

Visualize LEDR
```{r}
ledr_plot<-pr_wide %>% ggplot(aes(x=Temperature, y=LEDR, colour = Lifestage))+
  geom_point(position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic();ledr_plot

#remove those with 0 respiration as this is not biologically possible 
ledr_plot<-pr_wide %>% filter(LEDR>0)%>% ggplot(aes(x=Temperature, y=LEDR, colour = Lifestage))+
  geom_point(position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 2.5)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic();ledr_plot

ggsave("figures/p_r_rates/LEDR_dots.png", ledr_plot, width = 8, height = 4)

# view by means 
ledr_plot_2<-pr_wide %>% 
  group_by(Species, Lifestage, Temperature)%>%
  filter(LEDR>0)%>%
  summarise(mean=mean(LEDR, na.rm=TRUE), se=sd(LEDR, na.rm=TRUE)/sqrt(length(LEDR)))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.3), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.3))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 2)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic();ledr_plot_2

ggsave("figures/p_r_rates/LEDR_means.png", ledr_plot_2, width = 8, height = 4)
```

Visualize Pnet
```{r}
#plot  values
pnet_plot<-pr_wide %>% ggplot(aes(x=Temperature, y=Pnet, colour = Lifestage))+
  geom_point(aes(), position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  #geom_text(aes(label=sample_id),hjust=0, vjust=0)+
  xlab(expression(bold("Temperature °C"))) +
  ylim(0, 3.1)+
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic(); pnet_plot

ggsave("figures/p_r_rates/Pnet_dots.png", pnet_plot, width = 8, height = 4)

# view by means 
pnet_plot_2<-pr_wide %>% 
  group_by(Species, Lifestage, Temperature)%>%
  summarise(mean=mean(Pnet, na.rm=TRUE), se=sd(Pnet, na.rm=TRUE)/sqrt(length(Pnet)))%>%
  
  ggplot(aes(x=as.factor(Temperature), y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.3), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.3))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 3)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic();pnet_plot_2

ggsave("figures/p_r_rates/Pnet_means.png", pnet_plot_2, width = 8, height = 4)
```

Visualize P gross
```{r}
#plot  values
pgross_plot<-pr_wide %>% ggplot(aes(x=Temperature, y=Pgross, colour = Lifestage))+
  geom_point(aes(), position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  #geom_text(aes(label=sample_id),hjust=0, vjust=0)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic(); pgross_plot

ggsave("figures/p_r_rates/Pgross_dots.png", pgross_plot, width = 8, height = 4)

# view by means 
pgross_plot_2<-pr_wide %>% 
  group_by(Species, Lifestage, Temperature)%>%
  summarise(mean=mean(Pgross, na.rm=TRUE), se=sd(Pgross, na.rm=TRUE)/sqrt(length(Pgross)))%>%
  
  ggplot(aes(x=as.factor(Temperature), y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.3), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.3))+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  ylim(0, 4.1)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic();pgross_plot_2

ggsave("figures/p_r_rates/Pgross_means.png", pgross_plot_2, width = 8, height = 4)
```

Visualize PR ratio (filter to <6 to remove outliers). 
```{r}
#plot  values
pr_ratio_plot<-pr_wide %>% filter(LEDR>0) %>% filter(PR_ratio>0) %>% filter(PR_ratio<6)%>%
  ggplot(aes(x=Temperature, y=PR_ratio, colour = Lifestage))+
  geom_point(aes(), position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  #geom_text(aes(label=sample_id),hjust=0, vjust=0)+
  #ylim(0,5)+
  geom_hline(yintercept=1, linetype="dashed")+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold("P:R Ratio"))) +
  theme_classic(); pr_ratio_plot

ggsave("figures/p_r_rates/PR_ratio_dots.png", pr_ratio_plot, width = 8, height = 4)

# view by means 
pr_ratio_plot_2<-pr_wide %>% 
  filter(PR_ratio>0) %>%
  filter(PR_ratio<6)%>%
  filter(LEDR>0)%>%
  group_by(Species, Lifestage, Temperature)%>%
  summarise(mean=mean(PR_ratio, na.rm=TRUE), se=sd(PR_ratio, na.rm=TRUE)/sqrt(length(PR_ratio)))%>%
  
  ggplot(aes(x=as.factor(Temperature), y=mean, colour = Lifestage))+
  geom_point(aes(colour=Lifestage), position=position_dodge(0.3), size=2)+
  geom_errorbar(aes(ymin=(mean)-se, ymax=(mean)+se, group=Lifestage), color="gray", width=0.1, position=position_dodge(0.3))+
  geom_hline(yintercept=1, linetype="dashed")+
  geom_line(aes(group=Lifestage), linewidth=1.2, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold("P:R Ratio"))) +
  theme_classic();pr_ratio_plot_2

ggsave("figures/p_r_rates/PR_ratio_means.png", pr_ratio_plot_2, width = 8, height = 4)
```
These data are pretty noisy. 

# Plot P and R with labels for individual observation for QCing

```{r}
#plot  values
pnet_samples<-pr_wide %>% ggplot(aes(x=Lifestage, y=Pnet, colour = Lifestage))+
  geom_point(position=position_jitterdodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  geom_text(aes(label=sample_id),hjust=0, vjust=0, position=position_jitterdodge(0.3))+
  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic(); pnet_samples

ggsave("figures/p_r_rates/Pnet_samples.png", pnet_samples, width = 16, height = 12)
```

```{r}
#plot  values
pgross_samples<-pr_wide %>% ggplot(aes(x=Lifestage, y=Pgross, colour = Lifestage))+
  geom_point(position=position_jitterdodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  geom_text(aes(label=sample_id),hjust=0, vjust=0, position=position_jitterdodge(0.3))+
  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic(); pgross_samples

ggsave("figures/p_r_rates/Pgross_samples.png", pgross_samples, width = 16, height = 12)
```

```{r}
#plot  values
ledr_samples<-pr_wide %>% ggplot(aes(x=Lifestage, y=LEDR, colour = Lifestage))+
  geom_point(position=position_jitterdodge(0.3))+
  facet_wrap(~Species, nrow=1)+
  geom_text(aes(label=sample_id),hjust=0, vjust=0, position=position_jitterdodge(0.3))+
  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic(); ledr_samples

ggsave("figures/p_r_rates/LEDR_samples.png", ledr_samples, width = 16, height = 12)
```

# Statistical tests on each metric normalized to surface area 

Run ANOVAs on each metric 

Pgross
```{r}
model1<-aov(Pgross~Lifestage*Species*Temperature, data=pr_wide)
summary(model1)

emm<-emmeans(model1, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model1, ~Species)
pairs(emm)
```
Lifestage x species is significant. Adult POR higher than recruit. 

LEDR
```{r}
model2<-aov(LEDR~Lifestage*Species*Temperature, data=pr_wide)
summary(model2)

emm<-emmeans(model2, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model2, ~Species)
pairs(emm)

emm<-emmeans(model2, ~Temperature)
pairs(emm)
```
LEDR affected by temperature and lifestage x species. POR recruit less than adults. POC recruits less than adults. Rates are lower at 33°C than 30°C. 30°C and 27°C are not different. Temperature effects are only moderately significant. 

Pnet
```{r}
model3<-aov(Pnet~Lifestage*Species*Temperature, data=pr_wide)
summary(model3)

emm<-emmeans(model3, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model3, ~Temperature | Lifestage | Species)
pairs(emm)

emm<-emmeans(model3, ~Lifestage | Temperature | Species)
pairs(emm)
```
Pnet affected by lifestage x species and lifestage x temperature. POR adults higher than recruit. Adults are higher than recruits at 27°C and 30°C in POR only. Pnet is decreased at 33°C in adult POC but not recruits. 

PR Ratio
```{r}
model4<-aov(PR_ratio~Lifestage*Species*Temperature, data=pr_wide)
summary(model4)

emm<-emmeans(model4, ~Temperature)
pairs(emm)
```

# View metabolic rates in a regression by recruit size 

View by size. 

Pgross
```{r}
pgross_plot_recruit<-pr_wide %>% filter(Lifestage=="Recruit") %>% 
  ggplot(aes(x=surface.area, y=Pgross, colour = Temperature))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Temperature), se=FALSE)+
 # geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  xlab(expression(bold(paste("Recruit Surface Area (cm"^2, ")")))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic()+
  ylim(0, 3.5)+
  theme(legend.position="none"); pgross_plot_recruit

ggsave("figures/p_r_rates/Pgross_recruit_size.png", pgross_plot_recruit, width = 6, height = 4)
```

Not a strong relationship from what I can see. Interesting there is a positive relationship in POR but not the other species. 

LEDR
```{r}
ledr_plot_recruit<-pr_wide %>% filter(Lifestage=="Recruit") %>% filter(LEDR>0) %>%
  ggplot(aes(x=surface.area, y=LEDR, colour = Temperature))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  #geom_smooth(method="lm", aes(group=Temperature), se=FALSE)+
  geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  xlab(expression(bold(paste("Recruit Surface Area (cm"^2, ")")))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic()+
  ylim(0, 1.75)+
  theme(legend.position="none"); ledr_plot_recruit

ggsave("figures/p_r_rates/LEDR_recruit_size.png", ledr_plot_recruit, width = 6, height = 4)
```

Maybe a positive relationship in POR, no relationship in the other species. 

Pnet
```{r}
Pnet_plot_recruit<-pr_wide %>% filter(Lifestage=="Recruit") %>% 
  ggplot(aes(x=surface.area, y=Pnet, colour = Temperature))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  #geom_smooth(method="lm", aes(group=Temperature), se=FALSE)+
  geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  xlab(expression(bold(paste("Recruit Surface Area (cm"^2, ")")))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic()+
  ylim(0, 2.5)+
  theme(legend.position="none"); Pnet_plot_recruit

ggsave("figures/p_r_rates/Pnet_recruit_size.png", Pnet_plot_recruit, width = 6, height = 4)
```

No strong relationship. 

## Statistical tests 

Pgross by size 
```{r}
model4a<- pr_wide%>%
          filter(Lifestage=="Recruit")%>%
  
  aov(Pgross~surface.area*Species, data=.)

summary(model4a)
```
No effects on P gross. 

LEDR by size 
```{r}
model4b<- pr_wide%>%
          filter(LEDR>0)%>%
          filter(Lifestage=="Recruit")%>%
  
  aov(LEDR~surface.area*Species, data=.)

summary(model4b)
```
No effects on LEDR. 

Pnet by size 
```{r}
model4c<- pr_wide%>%
          filter(Lifestage=="Recruit")%>%
  
  aov(Pnet~surface.area*Species, data=.)

summary(model4c)
```
No effects on P net. 

# Correlate P gross and LEDR 

Plot correlation between P and R
```{r}
corr1<-pr_wide%>%
  filter(LEDR>0)%>%
  
  ggplot(aes(x=LEDR, y=Pgross, colour = Temperature))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  #geom_smooth(method="lm", aes(group=Temp.Cat.x), se=FALSE)+
  geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+
  scale_colour_manual(values=c("blue2", "orange", "red3"))+

  xlab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic()+
  ylim(0, 5.5)+
  theme(legend.position="none"); corr1

corr2<-pr_wide%>%
  filter(LEDR>0)%>%
  
  ggplot(aes(x=LEDR, y=Pgross, colour = Species))+
  geom_point(position=position_dodge(0.3), alpha=0.3)+
  geom_smooth(method="lm", aes(group=Species, colour=Species), se=FALSE)+
  scale_colour_manual(values=c("gray", "orange", "purple"))+

  xlab(expression(bold(paste("LEDR (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cm"^-2, "hr"^-1, ")")))) +
  theme_classic()+
  ylim(0, 5.5)+
  theme(legend.position="none"); corr2

ggsave("figures/p_r_rates/P_LEDR_correlation.png", corr1, width = 6, height = 4)
ggsave("figures/p_r_rates/P_LEDR_correlation_species.png", corr2, width = 6, height = 4)
```

There is a strong correlation between LEDR and Pgross. 
```{r}
model5<-pr_wide%>%
  filter(LEDR>0)%>%
  aov(Pgross~LEDR*Species, data=.)

summary(model5)

model5b<-pr_wide%>%
  filter(LEDR>0)%>%
  lm(Pgross~LEDR*Species, data=.)

coef(model5b)

cor.test(pr_wide$Pgross, pr_wide$LEDR, method="pearson")
```
The relationship between P and LEDR is strong. LEDR x species effect. Correlation is r=0.9128 and p<0.001. There is a different slope between species driven by POR. 

# View metrics normalized to symbiont cell density 

Plot metrics per cell. Identify outliers 
```{r}
dev.off()
hist(pr_wide$Pgross_cell)
hist(pr_wide$LEDR_cell)
hist(pr_wide$Pnet_cell)
hist(pr_wide$cells.cm2)
```

We have some outliers that we will need to remove for this analysis. 

First view symbiont cell density observations. 
```{r}
pr_wide%>%
  #filter(Pgross_cell<0.00003)%>%
  
  ggplot(aes(x=Temperature, y=cells.cm2, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Lifestage, colour=Lifestage), se=FALSE)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Cells cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right")
```
I don't see any obvious outliers. Keep all observations. 

## P gross 

View individual observations. 
```{r}
cell_plot1<-pr_wide%>%
  filter(Pgross_cell<0.00003)%>%
  
  ggplot(aes(x=Temperature, y=Pgross_cell, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); cell_plot1

ggsave("figures/p_r_rates/Pgross_cell_dots.png", cell_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
cell_plot2<-pr_wide%>%
  filter(Pgross_cell<0.00003)%>%
  filter(!is.na(Pgross_cell))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(Pgross_cell, na.rm=TRUE), sd=sd(Pgross_cell, na.rm=TRUE), N=length(Pgross_cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); cell_plot2

ggsave("figures/p_r_rates/Pgross_cell_means.png", cell_plot2, width = 6, height = 4)
```

Analyze Pgross per cell 

Run anova. 
```{r}
model6<-pr_wide%>%
  filter(Pgross_cell<0.00003)%>%
  filter(!is.na(Pgross_cell))%>%
  
  aov(Pgross_cell~Species*Lifestage*Temperature, data=.)

summary(model6)

emm<-emmeans(model6, ~Lifestage | Species)
pairs(emm)
```
Interaction between species x lifestage. Almost a 3-way interaction but not quite. Adults are lower than recruits in POC. Trend for higher per cell P at elevated temps in POC. 

## LEDR per cell 

Plot LEDR per cell 

View individual observations. 
```{r}
cell_plot3<-pr_wide%>%
  filter(LEDR_cell<0.000009)%>%
  
  ggplot(aes(x=Temperature, y=LEDR_cell, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); cell_plot3

ggsave("figures/p_r_rates/LEDR_cell_dots.png", cell_plot3, width = 6, height = 4)
```

View mean observations. 
```{r}
cell_plot4<-pr_wide%>%
  filter(LEDR_cell<0.000009)%>%
  filter(!is.na(LEDR_cell))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(LEDR_cell, na.rm=TRUE), sd=sd(LEDR_cell, na.rm=TRUE), N=length(LEDR_cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); cell_plot4

ggsave("figures/p_r_rates/LEDR_cell_means.png", cell_plot4, width = 6, height = 4)
```

Analyze LEDR per cell 

```{r}
model7<-pr_wide%>%
  filter(LEDR_cell<0.000009)%>%
  
  aov(LEDR_cell~Species*Lifestage*Temperature, data=.)

summary(model7)

emm<-emmeans(model7, ~Lifestage | Temperature | Species)
pairs(emm)

emm<-emmeans(model7, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model7, ~Temperature)
pairs(emm)
```
Lifestage x species x temperature effects. Lower LEDR per cell in adults than recruits in POC at 30°C. No other differences. Species x lifestage effects show xxxx. Temperature effects show higher values in recruits than adults of POC. LEDR per cell is higher at 30°C than 33°C in general. 

## P net per cell 

View individual observations. 
```{r}
cell_plot5<-pr_wide%>%
  filter(Pnet_cell<0.00002)%>%
  
  ggplot(aes(x=Temperature, y=Pnet_cell, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); cell_plot5

ggsave("figures/p_r_rates/Pnet_cell_dots.png", cell_plot5, width = 6, height = 4)
```

View mean observations. 
```{r}
cell_plot6<-pr_wide%>%
  filter(Pnet_cell<0.00002)%>%
  filter(!is.na(Pnet_cell))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(Pnet_cell, na.rm=TRUE), sd=sd(Pnet_cell, na.rm=TRUE), N=length(Pnet_cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " cell"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); cell_plot6

ggsave("figures/p_r_rates/Pnet_cell_means.png", cell_plot6, width = 6, height = 4)
```

Analyze Pnet per cell 

Run anova. 
```{r}
model8<-pr_wide%>%
  filter(Pnet_cell<0.00002)%>%
  filter(!is.na(Pnet_cell))%>%
  
  aov(Pnet_cell~Species*Lifestage*Temperature, data=.)

summary(model8)

emm<-emmeans(model8, ~Lifestage | Species)
pairs(emm)
```
Species x lifestage effects. Lower in adults than recruits of POC. Clear trend for higher Pnet per cell in POC at 30-33C, but temperature is not significant. Likely due to variability. 


# View metrics normalized to protein 

Plot metrics per mg protein. Identify outliers 
```{r}
dev.off()
hist(pr_wide$Pgross_prot)
hist(pr_wide$LEDR_prot)
hist(pr_wide$Pnet_prot)
hist(pr_wide$prot_mg.cm2)
```

We have some outliers that we will need to remove for this analysis. 

First view protein observations. 
```{r}
pr_wide%>%
  
  ggplot(aes(x=Temperature, y=prot_mg.cm2, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="lm", aes(group=Lifestage, colour=Lifestage), se=FALSE)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Protein (mg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right")
```
I don't see any obvious outliers. Keep all observations. 

We are missing POC-A11 and POR-R28. 

## P gross 

View individual observations. 
```{r}
prot_plot1<-pr_wide%>%
  filter(Pgross_prot<20)%>%
  
  ggplot(aes(x=Temperature, y=Pgross_prot, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot1

ggsave("figures/p_r_rates/Pgross_prot_dots.png", prot_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
prot_plot2<-pr_wide%>%
  filter(Pgross_prot<20)%>%
  filter(!is.na(Pgross_prot))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(Pgross_prot, na.rm=TRUE), sd=sd(Pgross_prot, na.rm=TRUE), N=length(Pgross_prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Gross], " (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); prot_plot2

ggsave("figures/p_r_rates/Pgross_prot_means.png", prot_plot2, width = 6, height = 4)
```

Analyze Pgross per mg prot 

Run anova. 
```{r}
model8<-pr_wide%>%
  filter(Pgross_prot<20)%>%
  filter(!is.na(Pgross_prot))%>%
  
  aov(Pgross_prot~Species*Lifestage*Temperature, data=.)

summary(model8)

emm<-emmeans(model8, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model8, ~Species)
pairs(emm)
```
Strong effects of lifestage and species. Adults are lower than recruits in ACR and POC, but posthoc shows no difference between adult and recruits in POR. At the species level, ACR and POC are not different, but POR is lower than the other two species.  

## LEDR per cell 

Plot LEDR per mg prot 

View individual observations. 
```{r}
prot_plot3<-pr_wide%>%
  filter(LEDR_prot<10)%>%
  
  ggplot(aes(x=Temperature, y=LEDR_prot, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); prot_plot3

ggsave("figures/p_r_rates/LEDR_prot_dots.png", prot_plot3, width = 6, height = 4)
```

View mean observations. 
```{r}
prot_plot4<-pr_wide%>%
  filter(LEDR_prot<10)%>%
  filter(!is.na(LEDR_prot))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(LEDR_prot, na.rm=TRUE), sd=sd(LEDR_prot, na.rm=TRUE), N=length(LEDR_prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("LEDR (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); prot_plot4

ggsave("figures/p_r_rates/LEDR_prot_means.png", prot_plot4, width = 6, height = 4)
```

Analyze LEDR per mg prot  

```{r}
model9<-pr_wide%>%
  filter(LEDR_prot<10)%>%
  
  aov(LEDR_prot~Species*Lifestage*Temperature, data=.)

summary(model9)

emm<-emmeans(model9, ~Lifestage | Temperature | Species)
pairs(emm)

emm<-emmeans(model9, ~Temperature | Species)
pairs(emm)

emm<-emmeans(model9, ~Temperature | Lifestage)
pairs(emm)

emm<-emmeans(model9, ~Temperature)
pairs(emm)
```
There are effects of species, lifestage, and temperature as well as species x temperature and lifestage x temperature. LEDR per mg protein is lower at 33 degrees than 30 degrees. There is a reduction in rates in ACR at 33C and increase from ambient to 30C in ACR but are stable in POR and POC when viewed by species as a whole. Recruits show a decrease at 33C whereas adults do not when looking at lifestage as a whole.  

At 30C, there are higher rates in recruits in POC but no temperature x lifestage differences in other species. 

## P net per mg prot 

View individual observations. 
```{r}
prot_plot5<-pr_wide%>%
  filter(Pnet_prot<20)%>%
  
  ggplot(aes(x=Temperature, y=Pnet_prot, colour = Lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~Species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=Lifestage), se=FALSE)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); prot_plot5

ggsave("figures/p_r_rates/Pnet_prot_dots.png", prot_plot5, width = 6, height = 4)
```

View mean observations. 
```{r}
prot_plot6<-pr_wide%>%
  filter(Pnet_prot<20)%>%
  filter(!is.na(Pnet_prot))%>%
  group_by(Species, Lifestage, Temperature)%>%
  mutate(mean=mean(Pnet_prot, na.rm=TRUE), sd=sd(Pnet_prot, na.rm=TRUE), N=length(Pnet_prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temperature, y=mean, colour = Lifestage, group=Lifestage))+
  geom_line(aes(group=Lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+
  facet_wrap(~Species, nrow=1, scales="free_x")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste(P[Net], " (µmol ", O[2], " mg prot"^-1, "hr"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="none"); prot_plot6

ggsave("figures/p_r_rates/Pnet_prot_means.png", prot_plot6, width = 6, height = 4)
```

Analyze Pnet per mg prot 

Run anova. 
```{r}
model10<-pr_wide%>%
  filter(Pnet_prot<20)%>%
  filter(!is.na(Pnet_prot))%>%
  
  aov(Pnet_prot~Species*Lifestage*Temperature, data=.)

summary(model10)

emm<-emmeans(model10, ~Lifestage | Species)
pairs(emm)

emm<-emmeans(model10, ~Lifestage)
pairs(emm)

emm<-emmeans(model10, ~Species)
pairs(emm)
```
Species and lifestage effects. Recruits are higher than adults in ACR and POC, but not in POR indicated by posthoc. Adults are lower than recruits on average. POR values are lower than ACR and POC on average.   

