---
title: "Chlorophyll analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")

# load packages
library(plotrix)
library(tidyverse)
library(ggplot2)
library(emmeans)

```

POR-A9 and ACR-A25 have 4 readings each, not 2. These are because samples were taken from both A and B tubes. 

90 observations - one is an extra or typo maybe? 


# Import data
```{r}
# Define function to read in chl data
read_chl <- function(file) {
  chl_data <- read_csv(file, skip = 24, n_max = 24) %>%
    select(-1) %>%
    magrittr::set_colnames(c("row", 1:12, "wavelength")) %>%
    fill(row) %>%
    gather("col", "absorbance", -wavelength, -row) %>%
    unite("well", c(row, col), sep = "")
}

# List chlorophyll data files
chl_path <- "data/physiology/chl/"                                        # Path to chlorophyll data directory
all_chl_files <- list.files(path = chl_path, pattern = "*.csv")          # List all files in directory
chl_platemaps <- list.files(path = chl_path, pattern = "platemap")       # List platemap files
chl_data_files <- setdiff(all_chl_files, chl_platemaps)                  # List absorbance data files

# Read in all files into tibble
df <- tibble(file = chl_data_files) %>%
  mutate(platemap = map(file, ~ read_csv(paste0(chl_path, tools::file_path_sans_ext(.), "_platemap.csv"))),
         chl_data = map(file, ~ read_chl(paste0(chl_path, .))))

# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, chl_data, ~ right_join(.x, .y)))
```

# Calculate chlorophyll concentrations
```{r}
# average all technical replicates for each plate/sample/wavelength, including all acetone blanks together (per plate)
df <- df %>%
  unnest(merged) %>%
  filter(!is.na(colony_id)) %>%                         # remove empty wells (colony_id is NA)
  group_by(file, colony_id, wavelength) %>%
  summarise(n = n(), mean_abs = mean(absorbance)) %>%
  spread(wavelength, mean_abs)

# get the acetone blank 750 absorbace for each file (i.e., plate), and subtract from 630 and 663 values for each sample
df <- df %>%
  group_by(file) %>%
  mutate(blank750 = `750`[colony_id == "BK"]) %>%
  ungroup() %>%
  mutate(adj630 = `630` - blank750,
         adj663 = `663` - blank750)

# calculate chla and chlc2 values based on equations from Jeffrey and Humphrey 1975
# units µg/ml
#path length adjustment = 0.6 

df <- df %>%
  mutate(chla.ug.ml = (11.43 * adj663)/0.6 - (0.64 * adj630)/0.6,
        chlc2.ug.ml = (27.09 * adj630)/0.6 - (3.63 * adj663)/0.6)

#previous, with no pathlength adjustment
#df <- df %>%
  #mutate(chla.ug.ml = (11.43 * adj663) - (0.64 * adj630),
        #chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663))
```

Remove any "B" replicates, these were taken from a duplicate slurry tube. All other metrics are taken only from the first replicate/A tube. 
```{r}
df<-df%>%
  filter(!str_detect(colony_id, "B"))

```

# Normalize to surface area
```{r}
# Load homogenate volume
homog.vol <- read_csv("data/physiology/homog_vols.csv") %>%
  select(sample_id, total_slurry_ml)
colnames(homog.vol)[1] <- "colony_id"

chl <- full_join(df, homog.vol)

# Load surface area
sa <- read_csv("output/physiology/surface_area/surface_area.csv")
colnames(sa)[1] <- "colony_id"
chl <- full_join(chl, sa)


# Multiply chlorophyll by the homogenate volume and divide by surface area
chl <- chl %>%
  mutate(chla.ug.cm2 = chla.ug.ml * total_slurry_ml / surface.area.cm2,
         chlc2.ug.cm2 = chlc2.ug.ml * total_slurry_ml / surface.area.cm2)

# remove blanks and NAs
chl <- filter(chl, !colony_id %in% c("NA", "BK"))

# write chlorophyll data to file
chl %>%
  select(colony_id, chla.ug.cm2, chlc2.ug.cm2) %>%
  filter(!is.na(chla.ug.cm2))%>%
  filter(!is.na(chlc2.ug.cm2))%>%
  write_csv(path = "output/chlorophyll.csv")
```

# Plot results by species and site
```{r, eval = TRUE}
# Join with sample metadata
metadata <- read_csv("data/field_collection/collection_metadata.csv") %>% select(sample_id, species, lifestage)
colnames(metadata)[1] <- "colony_id"
chl <- left_join(chl, metadata, by = "colony_id")

chl<-chl%>%filter(!is.na(file))

# Plot chlorophyll a
chl %>%
  ggplot(aes(x = lifestage, y = chla.ug.cm2, color = species)) +
  facet_wrap(~species) +
  labs(x = "", y = "chlorophyll a (µg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean

# Plot chlorophyll c2
chl %>%
  ggplot(aes(x = lifestage, y = chlc2.ug.cm2, color = species)) +
  facet_wrap(~species) +
  labs(x = "", y = "chlorophyll c2 (µg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean
```

No obvious outliers. 

# Plotting and ANOVA analyses 

Attach temperature metadata. 
```{r}
sample.info <- read_csv(file = "data/p_r_rates/pr_sample_metadata.csv")
sample.info<-sample.info%>%rename(colony_id=sample_id)%>%select(colony_id, Temp.Cat)%>%filter(!grepl("BK", colony_id))

chl<-right_join(chl, sample.info)
```

Check the number in each experimental group to find out which sample is the extra. 

```{r}
chl %>%
  group_by(lifestage, species, Temp.Cat)%>%
  summarise(N=length(chla.ug.cm2))
```
All are correct. 

Calculate total chlorophyll. 
```{r}
chl<-chl%>%
  mutate(total.ug.cm2=chla.ug.cm2+chlc2.ug.cm2)
```

Remove na's. None exist. 
```{r}
chl<-chl%>%
  filter(!is.na(file))
```

# Chl a normalized to surface area 

View individual observations. 
```{r}
chla_plot1<-chl%>%
  
  ggplot(aes(x=Temp.Cat, y=chla.ug.cm2, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Chlorophyll a (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chla_plot1

ggsave("figures/physiology/chlorophyll/chla_dots.png", chla_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
chla_plot2<-chl%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(chla.ug.cm2, na.rm=TRUE), sd=sd(chla.ug.cm2, na.rm=TRUE), N=length(chla.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Chlorophyll a (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chla_plot2

ggsave("figures/physiology/chlorophyll/chla_means.png", chla_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
chla_plot3<-chl%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(chla.ug.cm2, na.rm=TRUE), sd=sd(chla.ug.cm2, na.rm=TRUE), N=length(chla.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Chlorophyll a (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chla_plot3

ggsave("figures/physiology/chlorophyll/chla_overall_means.png", chla_plot3, width = 6, height = 4)
```

Plot data with sample labels. 

```{r}
sample_plot<-chl %>%

  ggplot(aes(x = lifestage, y = chla.ug.cm2, colour=lifestage)) +
  labs(x = "",y = "Chl a ug.cm2") +
  facet_wrap(~ species) +
  geom_jitter(aes(colour=lifestage),width = 0.1) +                                            # Plot all points
  theme_classic() + 
  geom_text(aes(label=colony_id),hjust=0, vjust=0, size=2);sample_plot

ggsave("figures/physiology/chlorophyll/chl_samples.png", sample_plot, dpi=300, w=6, h=4, units="in")
```

## Run ANOVA models

```{r}
model1<-chl%>%
  
  aov(chla.ug.cm2~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Species and lifestage and species x lifestage effects. Lower chl content in recruits of POC and POR relative to adults. No difference in ACR. Higher in POR overall. 


# Chl c2 normalized to surface area 

View individual observations. 
```{r}
chlc2_plot1<-chl%>%
  
  ggplot(aes(x=Temp.Cat, y=chlc2.ug.cm2, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Chlorophyll c2 (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chla_plot1

ggsave("figures/physiology/chlorophyll/chlc2_dots.png", chlc2_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
chlc2_plot2<-chl%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(chlc2.ug.cm2, na.rm=TRUE), sd=sd(chlc2.ug.cm2, na.rm=TRUE), N=length(chlc2.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Chlorophyll c2 (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chlc2_plot2

ggsave("figures/physiology/chlorophyll/chlc2_means.png", chlc2_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
chlc2_plot3<-chl%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(chlc2.ug.cm2, na.rm=TRUE), sd=sd(chlc2.ug.cm2, na.rm=TRUE), N=length(chlc2.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Chlorophyll c2 (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chlc2_plot3

ggsave("figures/physiology/chlorophyll/chla_overall_means.png", chlc2_plot3, width = 6, height = 4)
```

## Run ANOVA models

```{r}
model1<-chl%>%
  
  aov(chlc2.ug.cm2~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Species and lifestage and species x lifestage effects. Lower chl content in recruits of POC and POR relative to adults. No difference in ACR. Higher in POR overall. 






# Total chl normalized to surface area 

View individual observations. 
```{r}
chlt_plot1<-chl%>%
  
  ggplot(aes(x=Temp.Cat, y=total.ug.cm2, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chlt_plot1

ggsave("figures/physiology/chlorophyll/total_chl_dots.png", chlt_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
chlt_plot2<-chl%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(total.ug.cm2, na.rm=TRUE), sd=sd(total.ug.cm2, na.rm=TRUE), N=length(total.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chlt_plot2

ggsave("figures/physiology/chlorophyll/total_chl_means.png", chlt_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
chlt_plot3<-chl%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(total.ug.cm2, na.rm=TRUE), sd=sd(total.ug.cm2, na.rm=TRUE), N=length(total.ug.cm2), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cm"^-2, ")")))) +
  theme_classic()+
  theme(legend.position="right"); chlt_plot3

ggsave("figures/physiology/chlorophyll/total_chl_overall_means.png", chlt_plot3, width = 6, height = 4)
```

## Run ANOVA models

```{r}
model1<-chl%>%
  
  aov(total.ug.cm2~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Species and lifestage and species x lifestage effects. Lower chl content in recruits of POC and POR relative to adults. No difference in ACR. Higher in POR overall. 

# Normalize to cell density and total protein  

Read in cell density and total protein and merge. 
```{r}
cells<-read_csv(file="output/physiology/cell_density/symbiont_densities.csv")%>%select(sample_id, cells.cm2)
colnames(cells)[1] <- "colony_id"

chl<-left_join(chl, cells)

prot<-read_csv(file="output/physiology/protein/protein.csv")%>%select(colony_id, prot_mg.cm2)

chl<-left_join(chl, prot)
```

Calculate total chl normalized to cells and normalized to protein. 
```{r}
chl<-chl%>%
  mutate(total.ugchl.mg.prot=total.ug.cm2/prot_mg.cm2)%>%
  mutate(total.ugchl.cell=total.ug.cm2/cells.cm2)
```

# Total chl normalized to cell density 

View individual observations. 
```{r}
total_cell_plot1<-chl%>%
  
  ggplot(aes(x=Temp.Cat, y=total.ugchl.cell, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cell"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_cell_plot1

ggsave("figures/physiology/chlorophyll/total_chl_cell_dots.png", total_cell_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
total_cell_plot2<-chl%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(total.ugchl.cell, na.rm=TRUE), sd=sd(total.ugchl.cell, na.rm=TRUE), N=length(total.ugchl.cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cell"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_cell_plot2

ggsave("figures/physiology/chlorophyll/total_chl_cell_means.png", total_cell_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
total_cell_plot3<-chl%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(total.ugchl.cell, na.rm=TRUE), sd=sd(total.ugchl.cell, na.rm=TRUE), N=length(total.ugchl.cell), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Total chlorophyll (µg cell"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_cell_plot3

ggsave("figures/physiology/chlorophyll/total_chl_cell_overall_means.png", total_cell_plot3, width = 6, height = 4)
```

## Run ANOVA models

```{r}
model1<-chl%>%
  
  aov(total.ugchl.cell~species*lifestage*as.factor(Temp.Cat), data=.)

summary(model1)

emm<-emmeans(model1, ~ Temp.Cat | species)
pairs(emm)

emm<-emmeans(model1, ~ species*Temp.Cat)
pairs(emm)
```
Species x temperature effects. 

# Total chl normalized to total protein

View individual observations. 
```{r}
total_prot_plot1<-chl%>%
  
  ggplot(aes(x=Temp.Cat, y=total.ugchl.mg.prot, colour = lifestage))+
  geom_point(position=position_dodge(0.3), alpha=1)+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_smooth(method="loess", aes(group=lifestage), se=FALSE)+
  #geom_smooth(method="lm", aes(group=Species), se=FALSE, colour="black")+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_prot_plot1

ggsave("figures/physiology/chlorophyll/total_chl_prot_dots.png", total_prot_plot1, width = 6, height = 4)
```

View mean observations. 
```{r}
total_prot_plot2<-chl%>%
  group_by(species, lifestage, Temp.Cat)%>%
  mutate(mean=mean(total.ugchl.mg.prot, na.rm=TRUE), sd=sd(total.ugchl.mg.prot, na.rm=TRUE), N=length(total.ugchl.mg.prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=Temp.Cat, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Temperature °C"))) +
  ylab(expression(bold(paste("Total chlorophyll (mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_prot_plot2

ggsave("figures/physiology/chlorophyll/total_chl_prot_means.png", total_prot_plot2, width = 6, height = 4)
```

View species x lifestage level observations without temperature as there does not appear to be an effect.  
```{r}
total_prot_plot3<-chl%>%
  group_by(species, lifestage)%>%
  mutate(mean=mean(total.ugchl.mg.prot, na.rm=TRUE), sd=sd(total.ugchl.mg.prot, na.rm=TRUE), N=length(total.ugchl.mg.prot), se=sd/sqrt(N))%>%
  
  ggplot(aes(x=lifestage, y=mean, colour = lifestage, group=lifestage))+
  geom_line(aes(group=lifestage), colour="darkgray", linewidth=1, position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, position=position_dodge(0.3))+
  facet_wrap(~species, nrow=1, scales="free_x")+
  geom_point(position=position_dodge(0.3), alpha=1, size=3)+

  xlab(expression(bold("Lifestage"))) +
  ylab(expression(bold(paste("Total chlorophyll (mg protein"^-1, ")")))) +
  theme_classic()+
  theme(legend.position="right"); total_prot_plot3

ggsave("figures/physiology/chlorophyll/total_chl_prot_overall_means.png", total_prot_plot3, width = 6, height = 4)
```

## Run ANOVA models

```{r}
model1<-chl%>%
  
  aov(total.ugchl.mg.prot~species*lifestage*Temp.Cat, data=.)

summary(model1)

emm<-emmeans(model1, ~ lifestage | species)
pairs(emm)

emm<-emmeans(model1, ~ species*lifestage)
pairs(emm)
```
Species x lifestage effects. ACR recruit chl per unit protein are higher than adults. No difference in other species. Trend for ACR higher than other species. 

